A higher-order adaptive DPG method for compressible flow problems

### 2013-05-01 docs/Jesse/Rice talk/pres.tex

@@

% defines newblock as null, giving compile issues otherwise
\let\newblock\relax 
\institute[ICES]{Institute for Computational Engineering and Sciences}

\title[A DPG method for Compressible Flow]{A Discontinuous Petrov-Galerkin method for compressible flow problems}
\date{August 8, 2012}
\author[J. Chan, L. Demkowicz, R. Moser]{Jesse Chan \\ Supervisors: Leszek Demkowicz, Robert Moser \\ Committee: Todd Arbogast, Omar Ghattas, Venkat Raman}

@@

\section{Introduction/Literature review}
\frame{
\frametitle{Phenomena in compressible flow}
\begin{figure}
\centering
\subfigure[Shock wave]{\includegraphics[scale=.0704]{figs/compFlowPics/bullet_shock.png}}
\subfigure[Boundary layer]{\includegraphics[scale=.16]{figs/compFlowPics/boundary_layer.png}}
%\subfigure{\includegraphics[scale=.15]{figs/compFlowPics/capsule.png}}
\end{figure}
Compressible flow plays an important role in the aerospace and energy industries - transonic and supersonic aircraft, combustion engines, etc.

}
\frame{
\frametitle{Compressible Navier-Stokes equations}
\begin{columns}[c]
\begin{column}{.475\textwidth}
Numerical difficulties:
\begin{itemize}
\item{} Resolving solution features (sharp, localized viscous-scale phenomena)
\begin{itemize}
\item{} Shocks
\item{} Boundary layers - resolution needed for drag/load
\item{} Turbulence (non-localized)
\end{itemize}
\item{} Nonlinear convergence and uniqueness of solutions
\item{} Stability of numerical schemes
\begin{itemize}
\item{} Coarse/adaptive grids
\item{} Higher order
\end{itemize}
\end{itemize}
\end{column}
\begin{column}{.475\textwidth}
\begin{figure}
\centering
\includegraphics[scale = .7]{figs/compFlowPics/airfoil_mesh.png}
\end{figure}
\end{column}
\end{columns}
\vspace{.2cm}
\begin{center}
Idea: begin first with the model problem of convection-diffusion.
\end{center}
}
\frame{
\frametitle{Robustness: convection-diffusion as a model problem}
\vspace{-.5cm}
\[
\div \left(\beta u\right) - \epsilon \Delta u = f, \quad \text{on }\Omega \in \mathbb{R}^3
\]
\vspace{-.5cm}
\begin{columns}[c]
\begin{column}{.49\textwidth}
In 1D: $\beta u' - \epsilon u'' = f$. Standard continuous Galerkin variational formulation: solve
\[
b(u,v) = \ell(v), \quad u,v\in H^1_0([0,1])
\]
where
\begin{align*}
b(u,v) &= \int_\Omega -\beta uv' + \epsilon u'v' \\
\ell(v) &= \int_\Omega f v 
\end{align*}
\end{column}
\begin{column}{.49\textwidth}
\begin{figure}
\centering
\includegraphics[scale=.35]{figs/GalerkinOscTight.png}
\caption{Solution for $f=1$. Oscillations in the standard Galerkin method for large Peclet numbers ${\rm Pe} \coloneqq \frac{h}{\epsilon}$.}
\end{figure}
\end{column}
\end{columns}
}

\frame{
\frametitle{Streamline-upwind Petrov-Galerkin (SUPG)}
SUPG solves $b_{\rm SUPG}(u,v) = l_{\rm SUPG}(v)$, where
\begin{align*}
b_{\rm SUPG}(u,v) &= b(u,v)+ \sum_{K} \int_{K} \tau \LRp{L_{\rm adv} v} L u \\
l_{\rm SUPG}(v) &= \ell(v) + \sum_{K} \int_{K} \tau \LRp{L_{\rm adv}v} f.
\end{align*}
\vspace{-.5cm}
\begin{columns}[c]
\begin{column}{.44\textwidth}
\vspace{-.5cm}
\begin{itemize}
\item{} $Lu = \div\left(\beta u\right) - \epsilon \triangle u$, $L_{\rm adv}u = \div \left(\beta u\right)$, and $\tau$ is a parameter. 
%\item{} Recovers ``exact'' artificial diffusion for $f=0$.
\item{} Effective for $f\neq 0$. %``exact'' artificial diffusion.
\item{} \emph{Residual-based} stabilization. 
\end{itemize}
\end{column}
\begin{column}{.54\textwidth}
\begin{figure}
\centering
\includegraphics[scale=.38]{figs/SUPG.png}
%\caption{Exact, Bubnov-Galerkin, and SUPG solutions.}
\end{figure}
\end{column}
\end{columns}
}

\frame{
\begin{columns}
\begin{column}{.49\textwidth}
Can be interpreted as a \emph{Petrov-Galerkin} method,
\[
b\LRp{u,\tilde{v}_{i}} = \ell\LRp{\tilde{v}_{i}}, \quad \forall i = 1,\ldots,N-1,
\]
where the SUPG test function $\tilde{v}_{i}$ is defined elementwise as\footnotemark
\[
\tilde{v}_{i} = \phi_i(x) + \tau L_{\rm adv} \phi_i.  
\]
\end{column}
\begin{column}{.49\textwidth}
\begin{figure}
\centering
\subfigure{\includegraphics[scale=.16]{figs/SUPGtest.png}}
\caption{SUPG test function $v_i$.}
\end{figure}
\end{column}
\end{columns}
\footnotetext{\bibentry{SUPG}}
}

\section{The DPG Method}
\frame{
\frametitle{DPG: a minimum residual method via optimal testing}
Given a trial space $U$ and Hilbert test space $V$,
\[
b(u,v) = \ell(v), \quad u\in U, \quad v\in V
\]
This is equivalent to the operator equation posed in $V'$
\[
Bu = \ell
\]
if we identify $B:U\rightarrow V'$ and $\ell\in V'$ such that 
\begin{align*}
\langle Bu,v\rangle_V &\coloneqq b(u,v), \quad u\in U, v\in V,\\
\langle \ell,v\rangle_V &\coloneqq \ell(v), \quad v\in V.
\end{align*}
We seek to minimize the dual functional over $U_h \subset U$
\[
J(u_h) = \frac{1}{2}\|Bu_h-\ell\|_{V'}^2 \coloneqq\frac{1}{2} \sup_{v\in V\setminus\{0\}} \frac{| b(u_h,v)-\ell(v)|^2}{\nor{v}_V^2}.
\]
}

\begin{frame}
Let $R_V: V \to V'$ be the isometric Riesz map st.\
\[
\langle R_V v,\delta v\rangle_V \coloneqq(v, \delta v)_V, \quad \forall \delta v \in V.
\]
Then, our functional $J(u_h)$ is equal to
\begin{equation*}
\min_{u_h\in U_h} J(u_h) = \frac{1}{2}\left\|Bu_h-\ell\right\|_{V'}^2 =  \frac{1}{2}\left\|R_V^{-1}(Bu_h-\ell)\right\|_V^2.
\end{equation*}
First order optimality: G\^ateaux derivative is zero in all directions $\delta u \in U_h$
\begin{align*}
\left(R_V^{-1}(Bu_h-\ell),R_V^{-1}B\delta u\right)_V &= 0, \quad \forall \delta u \in U_h.\\
\rightarrow \LRa{(Bu_h-\ell),R_V^{-1}B\delta u} &= 0, \\
\rightarrow b(u_h,R_V^{-1}B\delta u) - \ell(R_V^{-1}B\delta u) &= 0
\end{align*}
\end{frame}

\frame{
\frametitle{Summary: select test functions to minimize residuals}
For $\delta u \in U_h$, define the \textcolor{red}{optimal test function} $v_{\delta u}$. 
\begin{equation*}
v_{\delta u} \coloneqq R_V^{-1}B\delta u.
\end{equation*} 
Then, the residual $J(u_h) = \frac{1}{2}\left\|Bu_h-\ell\right\|_{V'}^2$ is minimized by the solution of
\[
b\LRp{u_h,v_{\delta u}} = \ell\LRp{v_{\delta u}}, \quad \forall \delta u \in U_h.
\]
}

\frame{
\frametitle{Practical details of DPG}
Computation of $v_{\delta u} \coloneqq R_V^{-1}B\delta u$ is \textbf{global} and \textbf{infinite-dimensional}.
\begin{itemize}
\item By choosing a \textcolor{red}{broken} test space $V$ and \textcolor{red}{localizable} norm $\nor{v}_V^2 = \sum_K \nor{v}_{V(K)}^2$, test functions can be determined locally.
\item In practice, we use an \textcolor{red}{enriched space} $V_h \subset V$, where $\dim(V_h) > \dim(U_h)$ elementwise, and \textcolor{red}{optimal test functions} are approximated by computing $v_{\delta u} \coloneqq R_{V_h}^{-1}B\delta u \in V_h$ through\footnote{\bibentry{DPG2}}
\[
\LRp{v_{\delta u},\delta v}_V = b(\delta u,\delta v), \quad \delta u\in U_h, \quad \forall \delta v\in V_h
\]
Typically, if $U_h = \mathcal{P}^p(\mathbb{R}^n)$, $V_h = \mathcal{P}^{p+\triangle p}(\mathbb{R}^n)$, where $\triangle p \geq n$.\footnote{\bibentry{practicalDPG}} 
\end{itemize}
%\footnotetext{\bibentry{DPG2}}
}

\frame{
\frametitle{Properties of DPG}
DPG provides a symmetric positive-definite stiffness matrix. Let $\{\phi_j\}_{j=1}^m$ be a basis for $U_h$, and $\{v_i\}_{i=1}^n$ a basis for $V_h$, such that $n>m$. Then, for 
\begin{align*}
B_{ji} &= b(\phi_j,v_i),\\
l_i &= \ell(v_i),
\end{align*}
DPG solves the discrete system for degrees of freedom $u$
\begin{align*}
\LRp{B^T R_V^{-1}B} u  &= \LRp{B^T R_V^{-1}} l,
\end{align*}
where, under a localizable norm and discontinuous test functions, $R_V^{-1}$ is block diagonal.
}

\frame{
\frametitle{Properties of DPG}
Additional properties of DPG include\footnotemark
\begin{itemize}
\item DPG provides the best approximation in the \textcolor{red}{energy norm}
\[
\|u\|_E = \|Bu\|_{V'} = \sup_{\nor{v}_V=1} \left|b(u,v)\right|.
\]
\item The energy error is computable through the residual
\[
\nor{u-u_h}_E = \nor{B(u-u_h)}_{V'} = \nor{R_V^{-1}(l-Bu_h)}_V = \nor{e}_V
\]
where the \textcolor{red}{error representation function} $e$ is defined through $(e,\delta v)_V = \ell(v)-b(u_h,\delta v)$ for all $\delta v\in V$. 
\end{itemize}
\footnotetext{\bibentry{DPG3pub}}
}

%\frame{
%\frametitle{Duality of norms}
%Duality of norms: for any trial norm $\nor{u}_U$,
%\[
%\nor{u}_{U} = \sup_{v \in V}\frac{b\LRp{u,v}}{\nor{v}_{V,U}}, \quad \nor{v}_{V,U} = \sup_{u \in U}\frac%{b\LRp{u,v}}{\nor{u}_{U}}.
%\]
%Likewise, for any test norm $\nor{v}_V$, 
%\[
%\nor{v}_{V} = \sup_{u \in U}\frac{b\LRp{u,v}}{\nor{U}_{U,V}}, \quad \nor{u}_{U,V} = \sup_{v \in V}\frac%{b\LRp{u,v}}{\nor{v}_{V}}.
%\]
%
%}

\frame{
\frametitle{Ultra-weak formulation}
\vspace{-.75cm}
\begin{columns}
\begin{column}{.64\textwidth}
Given a first order system $Au = f$, we identify the \textbf{\emph{partition}} $\Oh$ and \textbf{\textcolor{red}{mesh skeleton}} \textcolor{red}{$\Gh$}.
\end{column}
\begin{column}{.34\textwidth}
%\[
%\Oh = \cup_{j=1}^\Nel K_j, \quad \Gh = \cup_{j=1}^\Nel \partial K_j
%\]
\begin{figure}[!h]
\centering
\includegraphics[scale = .155]{mesh_skel.pdf}
\end{figure}
\end{column}
\end{columns}
\vspace{-.5cm}
The ultra-weak formulation for $Au = f$ on $\Oh$ is
\[
b\left(\left(u, \widehat{u}\right),v\right) \coloneqq \sum_K\langle \widehat{u},v \rangle_{\partial K} + (u,A_h^*v)_{\Oh}= \LRp{f,v}_{\Oh}.
\]
Under proper assumptions, $\sum_K\langle \widehat{u},v \rangle_{\partial K} = \langle \widehat{u}, \jump{v} \rangle_{\Gh}$, with energy setting
\[
u\in L^2\LRp{\Oh} \equiv L^2(\Omega), \quad v\in V=D(A^*_h), \quad
\widehat{u}\in \gamma(D(A)),
\]
where $D(A_h^*)$ is the broken graph space of the formal adjoint $A_h^*$, and $\gamma(D(A))$ the trace space of the graph space of operator $A$.\footnote{\bibentry{analysisDPG}}
}

\frame{
\frametitle{The canonical ``graph'' test norm}
Recall $\nor{u}_E \coloneqq \sup_{v\in V\setminus\{0\}}\frac{b(u,v)}{\nor{v}_V}$. Under the ultra-weak formulation, the trial norm 
\[
\nor{\LRp{u,\widehat{u}}}_U \coloneqq \|u\|^2_{\L} + \|\widehat{u}\|^2
\] 
generates a test norm \emph{equivalent} to 
\[
\nor{v}_{V_{\rm opt}} \coloneqq \|A_h^*v\|_{\L}^2 +\left(\sup_{\widehat{u}} \frac{\LRa{ \widehat{u},\jump{v} }_{\Gh}}{\|\widehat{u}\|}\right)^2.
\]
This norm is not localizable, so we instead substitute the jump terms for an $L^2$ term, giving us the \emph{graph norm}\footnote{\bibentry{Bui-ThanhDemkowiczGhattas11b}}
\[
\nor{v}_V \coloneqq \|A_h^*v\|_{\L}^2 + \nor{v}_{L^2}.
\]
}

\frame{
\frametitle{Goal: robust higher order adaptive methods}
\begin{columns}
\begin{column}{.45\textwidth}
Our aim is to design an adaptive method for compressible laminar flows that is 
\begin{itemize}
\item robust in Reynolds number
\item stable for arbitrary elements
\end{itemize}

This proposal will outline
\begin{enumerate}
\item a robust DPG method for convection-diffusion,
\item extension to nonlinear problems and systems,
\item proposed work.
\end{enumerate}

\end{column}
\begin{column}{.54\textwidth}
\begin{figure}
\centering
\subfigure{\includegraphics[scale=.23]{figs/Re1000p2/u2_full.png}}
\subfigure{\includegraphics[scale=.37]{figs/Re1000p2/mesh6.png}}
\caption{$u_2$ and $p=2$ mesh for $\Reyn = 1000$.}
\end{figure}
\end{column}
\end{columns}
}

@@

\begin{column}{.49\textwidth}
\begin{figure}
\includegraphics[scale=.23]{figs/primal.png}
\caption{Non-standard inflow.}
\end{figure}
\end{column}
\end{columns}
When $\sigma_n \approx 0$ near the inflow, condition on $\widehat{f}_n$ approximates condition on $u$.

}

>>>>

\newcommand{\bibfoot}[1]{\footnote{\tiny\bibentry{#1}}}

@@

% defines newblock as null, giving compile issues otherwise
\let\newblock\relax

\institute[ICES]{Institute for Computational Engineering and Sciences}

\title[DPG methods for convective-diffusive problems]{A higher-order adaptive DPG method for compressible flow problems}
%\title[A DPG method for Compressible Flow]{Application of a Discontinuous Petrov-Galerkin method to problems in compressible flow}
\date{April 19, 2013}
\author[J. Chan, L. Demkowicz, R. Moser]{Jesse Chan\\ Supervisors: Leszek Demkowicz, Robert Moser}

@@

\section{Motivation}
\frame{
\frametitle{Phenomena in compressible flow}
\begin{figure}
\centering
\subfigure[Shock wave]{\includegraphics[scale=.0704]{figs/compFlowPics/bullet_shock.png}}
\subfigure[Boundary layer]{\includegraphics[scale=.16]{figs/compFlowPics/boundary_layer.png}}
%\subfigure{\includegraphics[scale=.15]{figs/compFlowPics/capsule.png}}
\end{figure}
Compressible flow plays an important role in the aerospace and energy industries - transonic and supersonic aircraft, combustion engines, etc. 
}

\frame{
\frametitle{Compressible Navier-Stokes equations}
\begin{columns}[c]
\begin{column}{.475\textwidth}
Numerical difficulties:
\begin{itemize}
\item{} Resolving solution features (sharp, localized viscous-scale phenomena)
\begin{itemize}
\item{} Shocks
\item{} Boundary layers - resolution needed for drag/load
\item{} Turbulence 
\end{itemize}
\item{} Nonlinear convergence and uniqueness of solutions
\item{} Stability of numerical schemes
\begin{itemize}
\item{} Coarse/adaptive grids
\item{} Higher order
\end{itemize}
\end{itemize}
\end{column}
\begin{column}{.475\textwidth}
\begin{figure}
\centering
\includegraphics[scale = .7]{figs/compFlowPics/airfoil_mesh.png}
\end{figure}
\end{column}
\end{columns}
\vspace{.2cm}
\begin{center}
Idea: begin first with the model problem of convection-diffusion.
\end{center}
}

\frame{
\frametitle{Robustness: convection-diffusion as a model problem}
\vspace{-.5cm}
\[
\div \left(\beta u\right) - \epsilon \Delta u = f, \quad \text{on }\Omega \in \mathbb{R}^3
\]
\vspace{-.5cm}
\begin{columns}[c]
\begin{column}{.49\textwidth}
In 1D: $\beta u' - \epsilon u'' = f$. Standard continuous Galerkin variational formulation: solve
\[
b(u,v) = \ell(v), \quad u,v\in H^1_0([0,1])
\]
where
\begin{align*}
b(u,v) &= \int_\Omega -\beta uv' + \epsilon u'v' \\
\ell(v) &= \int_\Omega f v 
\end{align*}
\end{column}
\begin{column}{.49\textwidth}
\begin{figure}
\centering
\includegraphics[scale=.35]{figs/GalerkinOscTight.png}
\caption{Solution for $f=1$. Oscillations in the standard Galerkin method for large Peclet numbers ${\rm Pe} \coloneqq \frac{h}{\epsilon}$.}
\end{figure}
\end{column}
\end{columns}
}

\frame{
\frametitle{Streamline-upwind Petrov-Galerkin (SUPG)}
SUPG solves $b_{\rm SUPG}(u,v) = l_{\rm SUPG}(v)$, where
\begin{align*}
b_{\rm SUPG}(u,v) &= b(u,v)+ \sum_{K} \int_{K} \tau \LRp{L_{\rm adv} v} L u \\
l_{\rm SUPG}(v) &= \ell(v) + \sum_{K} \int_{K} \tau \LRp{L_{\rm adv}v} f.
\end{align*}
\vspace{-.5cm}
\begin{columns}[c]
\begin{column}{.44\textwidth}
\vspace{-.5cm}
\begin{itemize}
\item{} $Lu = \div\left(\beta u\right) - \epsilon \triangle u$, $L_{\rm adv}u = \div \left(\beta u\right)$, and $\tau$ is a parameter. 
%\item{} Recovers ``exact'' artificial diffusion for $f=0$.
\item{} Restores discrete coercivity.
\item{} Effective for $f\neq 0$. %``exact'' artificial diffusion.
\item{} \textcolor{red}{Residual-based} stabilization. 
\end{itemize}
\end{column}
\begin{column}{.54\textwidth}
\begin{figure}
\centering
\includegraphics[scale=.38]{figs/SUPG.png}
%\caption{Exact, Bubnov-Galerkin, and SUPG solutions.}
\end{figure}
\end{column}
\end{columns}
}

\frame{
\begin{columns}
\begin{column}{.49\textwidth}
Can be ``interpreted'' as a \textcolor{red}{Petrov-Galerkin} method,
\[
b\LRp{u,\tilde{v}_{i}} = \ell\LRp{\tilde{v}_{i}}, \quad \forall i = 1,\ldots,N-1,
\]
where the SUPG test function $\tilde{v}_{i}$ is defined elementwise as\footnotemark
\[
\tilde{v}_{i} = \phi_i(x) + \tau L_{\rm adv} \phi_i.  
\]
\end{column}
\begin{column}{.49\textwidth}
\begin{figure}
\centering
\subfigure{\includegraphics[scale=.16]{figs/SUPGtest.png}}
\caption{SUPG test function $v_i$.}
\end{figure}
\end{column}
\end{columns}
\footnotetext{\tiny\bibentry{SUPG}}
}

%\section{The DPG Method}
%\frame{
%\frametitle{DPG: a minimum residual method via optimal testing}
%Given a trial space $U$ and Hilbert test space $V$,
%\[
%b(u,v) = \ell(v) \Longleftrightarrow Bu = \ell, \quad 
%\left.\begin{array}{cl}
%\langle Bu,v\rangle_V &\coloneqq b(u,v)\\
%\langle \ell,v\rangle_V &\coloneqq \ell(v).
%\end{array}\right.
%\]
%We seek to minimize the \textcolor{red}{dual residual} over $U_h \subset U$
%\[
%J(u_h) = \frac{1}{2}\|Bu_h-\ell\|_{V'}^2 = \frac{1}{2}\left\|R_V^{-1}(Bu_h-\ell)\right\|_V^2.
%\]
%where $R_V: V \to V'$ is the isometric Riesz map
%\[
%\langle R_V v,\delta v\rangle_V \coloneqq(v, \delta v)_V, \quad \forall \delta v \in V.
%\]
%Conditions for minimization of the convex functional give
%\[
%b(u_h,v_{\delta u}) = \ell(v_{\delta u}), \quad \forall \delta u \in U_h, \quad v_{\delta u} \coloneqq R_V^{-1}B\delta u.
%\]
%}

\section{The DPG Method}
\frame{
\frametitle{DPG: a minimum residual method via optimal testing}
Given a trial space $U$ and Hilbert test space $V$,
\[
b(u,v) = \ell(v), \quad u\in U, \quad v\in V
\]
This is equivalent to the operator equation posed in $V'$
\[
Bu = \ell
\]
if we identify $B:U\rightarrow V'$ and $\ell\in V'$ such that 
\begin{align*}
\langle Bu,v\rangle_V &\coloneqq b(u,v), \quad u\in U, v\in V,\\
\langle \ell,v\rangle_V &\coloneqq \ell(v), \quad v\in V.
\end{align*}
We seek to minimize the \textcolor{red}{dual residual} over $U_h \subset U$
\[
J(u_h) = \frac{1}{2}\|Bu_h-\ell\|_{V'}^2 \coloneqq\frac{1}{2} \sup_{v\in V\setminus\{0\}} \frac{| b(u_h,v)-\ell(v)|^2}{\nor{v}_V^2}.
\]
}

\begin{frame}
Let $R_V: V \to V'$ be the isometric Riesz map st.\
\[
\langle R_V v,\delta v\rangle_V \coloneqq(v, \delta v)_V, \quad \forall \delta v \in V.
\]
Then, our residual $J(u_h)$ is equal to
\begin{equation*}
\min_{u_h\in U_h} J(u_h) = \frac{1}{2}\left\|Bu_h-\ell\right\|_{V'}^2 =  \frac{1}{2}\left\|R_V^{-1}(Bu_h-\ell)\right\|_V^2.
\end{equation*}
First order optimality: G\^ateaux derivative is zero in all directions $\delta u \in U_h$
\begin{align*}
\left(R_V^{-1}(Bu_h-\ell),R_V^{-1}B\delta u\right)_V &= 0, \quad \forall \delta u \in U_h.\\
\rightarrow \LRa{(Bu_h-\ell),R_V^{-1}B\delta u} &= 0, \\
\rightarrow b(u_h,R_V^{-1}B\delta u) - \ell(R_V^{-1}B\delta u) &= 0
\end{align*}
\end{frame}

\frame{
\frametitle{Summary: select test functions to minimize residuals}
For $\delta u \in U_h$, define \textcolor{red}{optimal test functions} $v_{\delta u}\coloneqq R_V^{-1}B\delta u$ as solutions to
\[
\LRp{v_{\delta u},\delta v}_V = b({\delta u},\delta v), \quad \forall \delta v \in V.
\]
The \textcolor{red}{optimal test space} $V_{\rm opt}$
\[
V_{\rm opt} \coloneqq \LRc{v_{\delta u}: \delta u\in U_h}.
\]
Then, the residual $J(u_h) = \frac{1}{2}\left\|Bu_h-\ell\right\|_{V'}^2$ is minimized by the solution of
\[
b\LRp{u_h,v} = \ell\LRp{v}, \quad \forall v \in V_{\rm opt}.
\]
}

\frame{
\frametitle{Practical details of DPG}
Computation of $v_{\delta u} \coloneqq R_V^{-1}B\delta u$ is \textbf{global} and \textbf{infinite-dimensional}.
\begin{itemize}
\item By choosing a \textcolor{red}{broken} test space $V$ and \textcolor{red}{localizable} norm $\nor{v}_V^2 = \sum_K \nor{v}_{V(K)}^2$, test functions can be determined locally.
\item In practice, we use an \textcolor{red}{enriched space} $V_h \subset V$, where $\dim(V_h) > \dim(U_h)$ elementwise, and \textcolor{red}{optimal test functions} are approximated by computing $v_{\delta u} \coloneqq R_{V_h}^{-1}B\delta u \in V_h$ through{\bibfoot{DPG2}}
\[
\LRp{v_{\delta u},\delta v}_V = b(\delta u,\delta v), \quad \delta u\in U_h, \quad \forall \delta v\in V_h
\]
Typically, if $U_h = \mathcal{P}^p(\mathbb{R}^n)$, $V_h = \mathcal{P}^{p+\triangle p}(\mathbb{R}^n)$, where $\triangle p \geq n$.{\bibfoot{practicalDPG}} 
\end{itemize}
}

\frame{
\frametitle{Properties of DPG}
DPG provides a \textcolor{red}{symmetric positive-definite} stiffness matrix. Let $\{\phi_j\}_{j=1}^m$ be a basis for $U_h$, and $\{v_i\}_{i=1}^n$ a basis for $V_h$, such that $n>m$. Then, for 
\begin{align*}
B_{ij} &= b(\phi_j,v_i),\\
l_i &= \ell(v_i),
\end{align*}
DPG solves the discrete system for degrees of freedom $u$
\begin{align*}
\LRp{B^T R_V^{-1}B} u  &= \LRp{B^T R_V^{-1}} l,
\end{align*}
where, under a localizable norm and discontinuous test functions, $R_V^{-1}$ is block diagonal.
}

\frame{
\frametitle{Properties of DPG}
Additional properties of DPG include\footnotemark
\begin{itemize}
\item DPG provides the best approximation in the \textcolor{red}{energy norm}
\[
\|u\|_E = \|Bu\|_{V'} = \sup_{\nor{v}_V=1} \left|b(u,v)\right|.
\]
\item The energy error is computable through the residual
\[
\nor{u-u_h}_E = \nor{B(u-u_h)}_{V'} = \nor{R_V^{-1}(l-Bu_h)}_V = \nor{e}_V
\]
where the \textcolor{red}{error representation function} $e$ is defined through $(e,\delta v)_V = \ell(v)-b(u_h,\delta v)$ for all $\delta v\in V$. 
\item Adaptivity driven by local error indicator $\nor{e}_{V(K)}^2$
\end{itemize}
\footnotetext{\tiny\bibentry{DPG3pub}}
}

%\frame{
%\frametitle{Alternative interpretations of DPG}
%DPG has been referred to as ``least-squares (with a twist)".  Recall that, given trial and test spaces $U$ and $V$, least-squares orthogonality of error is
%\begin{align*}
%\min_{u_h\in U_h}\frac{1}{2}\|Bu_h-\ell\|_{V'}^2 \Rightarrow \left(R_V^{-1}(Bu_h-\ell),R_V^{-1}B\delta u\right)_V = 0, \quad \forall \delta u \in U_h.
%\end{align*}
%By undoing the Riesz operator on the \textcolor{red}{left}, we arrive at optimal test functions 
%\begin{align*}
%\LRa{Bu_h-\ell,R_V^{-1}B\delta u} = b(u,R_V^{-1}B\delta u)-l(R_V^{-1}B\delta u) = 0, \quad \forall \delta u \in U_h
%\end{align*}
%}

\frame{
\frametitle{Residual-based stabilization (with a twist)}
Can rewrite least-squares minimization as a \textcolor{red}{saddle point problem}.\bibfoot{DahmenVariationalStabilization}
%\begin{align*}
%\min_{u_h\in U_h}\frac{1}{2}\|Bu_h-\ell\|_{V'}^2 \Rightarrow \left(R_V^{-1}(Bu_h-\ell),R_V^{-1}B\delta u\right)_V = 0, \quad \forall \delta u \in U_h.
%\end{align*}
Define 
$$R_V^{-1}(Bu-l)  = e \in V.$$ 
Undoing the \textcolor{red}{right} Riesz operator leads to the Galerkin orthogonality condition
\[
\LRp{e,R_V^{-1}B\delta u}_V = \LRa{e,B\delta u} = b(\delta u, e) = 0, \quad \forall \delta u \in U_h
\]
which gives the system for $\LRp{u,e} \in U\times V$
\begin{align*}
b(u,v) + \LRp{e,v}_V &= l(v), \quad \forall v \in V\\
b(\delta u, e) &= 0, \quad \forall \delta u \in U_h
\end{align*}
\vspace{-.2cm}
%\begin{itemize}
%\item $\LRp{e,v}_V$ can be interpreted as a VMS-type contribution.
%\item Can be extended to arbitrary variational formulations.
%\end{itemize}
}

\frame{
\frametitle{Ultra-weak formulation}
\vspace{-.75cm}
\begin{columns}
\begin{column}{.64\textwidth}
Given a first order system $Au = f$, we identify the \textbf{\emph{partition}} $\Oh$ and \textbf{\textcolor{red}{mesh skeleton}} \textcolor{red}{$\Gh$}.
\end{column}
\begin{column}{.34\textwidth}
%\[
%\Oh = \cup_{j=1}^\Nel K_j, \quad \Gh = \cup_{j=1}^\Nel \partial K_j
%\]
\begin{figure}[!h]
\centering
\includegraphics[scale = .155]{figs/mesh_skel.pdf}
\end{figure}
\end{column}
\end{columns}
\vspace{-.5cm}
The ultra-weak formulation for $Au = f$ on $\Oh$ is
\[
b\left(\left(u, \widehat{u}\right),v\right) \coloneqq \sum_K\langle \widehat{u},v \rangle_{\partial K} + (u,A_h^*v)_{\Oh}= \LRp{f,v}_{\Oh}.
\]
Under proper assumptions, $\sum_K\langle \widehat{u},v \rangle_{\partial K} = \langle \widehat{u}, \jump{v} \rangle_{\Gh}$, with energy setting
\[
u\in L^2\LRp{\Oh} \equiv L^2(\Omega), \quad v\in V=D(A^*_h), \quad
\widehat{u}\in \gamma(D(A)),
\]
where $D(A_h^*)$ is the broken graph space of the formal adjoint $A_h^*$, and $\gamma(D(A))$ the trace space of the graph space of operator $A$.\bibfoot{analysisDPG}
}

\section{Test norms for DPG}

\frame{
\frametitle{The canonical ``graph'' test norm}
Recall $\nor{u}_E \coloneqq \sup_{v\in V\setminus\{0\}}\frac{b(u,v)}{\nor{v}_V}$. Under the ultra-weak formulation, the trial norm 
\[
\nor{\LRp{u,\widehat{u}}}_U^2 \coloneqq \|u\|^2_{\L} + \|\widehat{u}\|^2
\] 
generates a test norm \emph{equivalent} to 
\[
\|A_h^*v\|_{\L}^2 +\left(\sup_{\widehat{u}} \frac{\LRa{ \widehat{u},\jump{v} }_{\Gh}}{\|\widehat{u}\|}\right)^2.
\]
This norm is not localizable, so we instead substitute the jump terms for an $L^2$ term, giving us the \textcolor{red}{graph norm}{\bibfoot{Bui-ThanhDemkowiczGhattas11b}}
\[
\nor{v}_V^2 \coloneqq \|A_h^*v\|_{\L}^2 + \nor{v}_{L^2}^2.
\]
}

\setcounter{subfigure}{0}
\frame{
\frametitle{The graph test norm - near $L^2$ best approx for Stokes.}
\vspace{-.3cm}
\begin{figure}
\centering
\subfigure[Manufactured solution:\ $u_1$ error]{\includegraphics[scale=.3]{figs/stokes/u1_err.pdf}}
\subfigure[Manufactured solution:\ $p$ error]{\includegraphics[scale=.3]{figs/stokes/pressure_err.pdf}}
\caption{Results under the graph norm for Stokes flow.{\bibfoot{stokesDPG}}}
\end{figure}
}
\frame{
\frametitle{The graph test norm - Stokes lid-driven cavity flow.}
\vspace{-.3cm}
\begin{figure}
\centering
\subfigure[Cavity flow streamlines]{\includegraphics[scale=.3]{figs/stokes/streamlines.pdf}}
\subfigure[Second Moffat eddy]{\includegraphics[scale=.3]{figs/stokes/streamlines_zoom.pdf}}
\end{figure}
\vspace{-.3cm}
Similar results for elasticity, electromagnetics, and wave propagation{\bibfoot{DPG4}}.
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\frame{
\frametitle{DPG as a non-conforming DG method on $V$}
For DPG using the ultra-weak variational formulation
\[
b\left(\left(u, \widehat{u}\right),v\right) \coloneqq \LRa{\widehat{u},\jump{v}}_{\Gh} + (u,A_h^*v)_{\Oh}= \LRp{f,v}_{\Oh},
\]
the saddle point problem is
\begin{align*}
\LRp{e,v}_V &= l(v)-b\LRp{\LRp{u,\uh},v}, \quad \forall v \in V(\Oh)\\
(\delta u, A^*_he)_{\L} &= 0, \quad \forall (\delta u, 0) \in U_h\\
\LRa{\widehat{\delta u},\jump{e}}_{\Gh} &= \int_{\Gh} \widehat{\delta u} \jump{e} =  0, \quad \forall \LRp{0,\widehat{\delta u}} \in U_h,
\end{align*}
This implies $e\in \tilde{V}$, the \emph{weakly conforming} space 
\[
\tilde{V} = \LRc{v \in V(\Oh): \LRa{\uh,\jump{v}}_{\Gh} = 0, \quad \forall \LRp{0,\widehat{\delta u}} \in U_h}.
\]
}

\frame{
\frametitle{Global properties of DPG test spaces}
If we define $\tilde{V}_{\rm opt}$ as the optimal test space constructed using $\tilde{V}$
\[
\tilde{V}_{\rm opt} = \LRc{ R_{\tilde V}^{-1}B \delta u, \quad \forall \delta u \in U_h},
\]
we have the following theorem\bibfoot{globalDPG}  
\begin{theorem}
$\tilde{V}_{\rm opt} \subset V_{\rm opt}$.  Furthermore, the $L^2$/interior solutions of $b(u,v)=l(v)$ over $V_{\rm opt}$ and $\tilde{V}_{\rm opt}$ coincide.
\end{theorem}
\begin{itemize}
\item The locally determined optimal test functions approximate globally optimal test functions.
\item Under a variant of the graph norm, DPG approximates test functions which deliver $L^2$ optimality.
\end{itemize}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{column}{.49\textwidth}
\begin{figure}
\includegraphics[scale=.23]{figs/primal.png}
\caption{Non-standard inflow.}
\end{figure}
\end{column}
\end{columns}
When $\sigma_n \approx 0$ near the inflow, condition on $\widehat{f}_n$ approximates condition on $u$. 
}
